---
title: "final+project"
output: html_document
date: "2025-04-29"
---
load libraries and data
```{r}
library(readr)
library(dbplyr)
library(tidyverse)
library(ggplot2)
library(shiny)
dfdirty <- read_csv("C:/Users/btaus/DC_Public_Schools_Student_Assessment_Results - Copy.csv")
```

Data Cleaning
```{r}
dfdirty <- read_csv("C:/Users/btaus/DC_Public_Schools_Student_Assessment_Results - Copy.csv")

colnames(dfdirty) <- gsub("[^[:alnum:]_]", "", colnames(dfdirty))

dfdirty <- as.data.frame(dfdirty)



selected_subgroups <- c("Female", "Male", "Black/African American", "White", "Hispanic/Latino", "Economically Disadvantaged")

dffinal <- dfdirty[dfdirty$SCHOOL_YEAR == "2014-15" & dfdirty$SUBJECT == "Math" & dfdirty$TESTED_GRADE_SUBJECT == "Grade 3" & dfdirty$ASSESSMENT_TYPE == "PARCC" , ] %>% 
  dplyr::select(SCHOOL_WARD, SCHOOL_NAME, ASSESSMENT_TYPE, SUBJECT, TESTED_GRADE_SUBJECT, SUBGROUP, SUBGROUP_VALUE, PERCENT_MEETING_OR_EXCEEDING_E, TOTAL_NUMBER_VALID_TEST_TAKERS, PERCENT_LEVEL_3_, PERCENT_LEVEL_1, PERCENT_LEVEL_2, PERCENT_LEVEL_3, PERCENT_LEVEL_4, PERCENT_LEVEL_5) %>%
  mutate(PERCENT_MEETING_OR_EXCEEDING_E = as.character(PERCENT_MEETING_OR_EXCEEDING_E),
         PERCENT_MEETING_OR_EXCEEDING_E = ifelse(PERCENT_MEETING_OR_EXCEEDING_E == "n<25", NA, PERCENT_MEETING_OR_EXCEEDING_E),
         PERCENT_MEETING_OR_EXCEEDING_E = gsub("%", "", PERCENT_MEETING_OR_EXCEEDING_E), 
         PERCENT_MEETING_OR_EXCEEDING_E = as.numeric(PERCENT_MEETING_OR_EXCEEDING_E)) %>% dplyr::select(SCHOOL_WARD, SCHOOL_NAME, ASSESSMENT_TYPE, SUBJECT, TESTED_GRADE_SUBJECT, SUBGROUP, SUBGROUP_VALUE, PERCENT_MEETING_OR_EXCEEDING_E, TOTAL_NUMBER_VALID_TEST_TAKERS) %>% 
  filter(SUBGROUP_VALUE %in% selected_subgroups)



write.csv(dffinal, "dffinal.csv")

```



Introduction: State your scientific question clearly in plain English and explain why it is interesting.

Is the number of exceeding scores between wards significantly different? how about within specific subgroups?

Exploratory Data Analysis (EDA):

Provide summary statistics and visuals.

```{r}
summary(dffinal)
```

```{r}
ggplot(dffinal, aes(x = PERCENT_MEETING_OR_EXCEEDING_E)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
  facet_wrap(~ SUBGROUP_VALUE, scales = "free_y") +
  labs(title = "Distribution of Test Scores by Subgroup",
       x = "Percent Meeting or Exceeding Expectations",
       y = "Count") +
  theme_minimal()
```

```{r}
ggplot(dffinal, aes(x = PERCENT_MEETING_OR_EXCEEDING_E)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
  facet_wrap(~ SCHOOL_WARD, scales = "free_y") +
  labs(title = "Distribution of Test Scores by Ward",
       x = "Percent Meeting or Exceeding Expectations",
       y = "Number of Schools") +
  theme_minimal()
```

```{r}
ggplot(dffinal, aes(x = SCHOOL_WARD, y = PERCENT_MEETING_OR_EXCEEDING_E, fill = SUBGROUP_VALUE)) +
  geom_boxplot(position = "dodge") +
  labs(title = "Test Scores by Ward and Subgroup",
       y = "Percent Meeting or Exceeding Expectations") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
ggplot(dffinal, aes(x = SCHOOL_WARD, y = PERCENT_MEETING_OR_EXCEEDING_E)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Test Score Distribution by Ward",
       y = "Percent Meeting or Exceeding Expectations") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```




Informally address your question.

Statistical Hypothesis:

Transform your scientific question into a formal statistical hypothesis.

Use notation where appropriate.

Statistical Analysis:

Implement and describe your analysis.

Include graphics to check model assumptions.

Conclusion:

Summarize your findings.

shiny code
```{r}
ui <- fluidPage(
  titlePanel("Test Score Distribution by Ward and Subgroup"),

  sidebarLayout(
    sidebarPanel(
      selectInput("subgroup", "Select Subgroup:",
                  choices = c("Female", "Male", "Black/African American", "White", "Hispanic/Latino", "Economically Disadvantaged"),
                  selected = "Female")
    ),
    mainPanel(
      plotOutput("distPlot")
    )
  )
)

server <- function(input, output) {

  filtered_data <- reactive({
    req(input$subgroup)
    dffinal %>%
      filter(SUBGROUP_VALUE == input$subgroup)
  })

  output$distPlot <- renderPlot({
    ggplot(filtered_data(), aes(x = PERCENT_MEETING_OR_EXCEEDING_E)) +
      geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
      facet_wrap(~ SCHOOL_WARD, scales = "free_y") +
      labs(title = paste("Distribution of Scores by Ward -", input$subgroup),
           x = "Percent Meeting or Exceeding Expectations",
           y = "Number of Schools") +
      theme_minimal()
  })
}

shinyApp(ui = ui, server = server)
```

